Kaggle-in-class Data Challenges Can Boost Student Learning

### 2018-12-20 submission1/response-to-review.txt

Reviewer(s)' Comments to Author:

Reviewer: 1

Comments to the Author
I enjoyed this paper and think it makes a contribution to the literature.  However,  I would like to see some improvements made.
Briefly discussing the experience from the professor side of things would be helpful.  So in addition to the student survey it would be helpful to read how hard it was to set up and work with the Kaggle competitions and any logistical challenges.

Also more discussion of study limitations is needed.  For example,  in the ETC class the "treatment" is confounded with level of student ( graduate/ undergraduate).  Other limitations should be briefly discussed as well.  I also would like to see suggestions for future research.
Sentence misworded in line 29 page 9
Also in Figure 3, it would seem there is some non-linearity in a couple of the graphs.  I would like to see that mentioned and discussed.  It seems there is a point of diminishing return, in graphs E and F.

I also think it is a study weakness that all the survey questions are worded in a positive manner and this could be leading.  I realize given the study is done this can’t be changed but it would be worth noting in the discussion ways the survey might be improved.
It might be more productive in future work to directly ask for level of engagement first, then one could also ask students who did not do the competition as well the question.
“How would you rate your level of engagement in this course?”
Not at all engaged--- up to extremely engaged with choices in between.

Reviewer: 2

Comments to the Author
Kaggle is an online platform that brings together two communities: one that is willing to contribute data and another that is interested in analyzing data. Kaggle has become one of the popular platforms used in quantitative statistical courses, where individual students or teams can obtain a data set to analyze, and often compete (for monetary prize or otherwise). The main aim of the article is to provide empirical evidence in support of the effectiveness of Kaggle competitions in improving student learning.
In terms of grammar, style, and narrative the article is written adequately. It gives a concise summary of the empirical results that it uses to support the claims of effectiveness. However, there are major issues related to the motivation and relatedly the (absence of) literature review; details and guidelines; and the quantitative assessment of effectiveness of the competition. The major issues are outlined below, followed by a few minor comments. In reviewer’s opinion, the adequate resolution of all of those observations is a necessary precondition for the publication of the work.

Major Comments
- The article does not provide a motivation for incorporating a predictive competition in a coursework. What is the educational aim of the competitive element as compared to more “standard” projects where students do not compete? Relatedly, what are the advantages and limitations of the proposed competitive format? A literature review, which currently is completely missing from the article, would be more than appropriate, summarizing thoroughly the existing work on project-based learning and putting the proposed format of Kaggle competitions vis-à-vis the project-based pedagogy.
- The article does not provide details regarding the implementation/administration of the competition in a classroom setting, and lacks practical guidelines for teachers of statistics and data science. Thus, in the current form it has little to offer if the reader considers replicating Kaggle competitions in her/his classroom. Some of the important details that are left out of the article include but are not limited to the time and involvement required from the instructor from start to finish; the format of instructor’s involvement with students/teams; time in the semester (start? middle? end?) when the projects get assigned; presence of a teaching assistant (TA), etc.
- Many of the datasets on Kaggle are in need of preprocessing/cleaning prior to the application of predictive models. In authors’ view, should these tasks be carried out by instructors/TAs or left to students instead? What are the pros and cons of each of these approaches?
- For many Kaggle datasets there are existing (at least preliminary) solutions on the web. Reliance on those resources may drastically deteriorate student involvement, interest, and takeaways from the projects. In authors’ view, what steps can be taken by instructors in that regard?
- In the first paragraph of section 2.3 authors mention that students were allowed to first submit individually and then form groups. It would be helpful if authors could explain the motivation behind that format.
- In the second paragraph of section 3.1, as well as in section 4.1 authors note/imply that the postgrad cohort in ETC2420 took part in the Kaggle challenge while the undergraduate cohort was used as the control group. This could introduce confounding, as the average quantitative preparedness in the postgraduate cohort could be significantly higher than that of the undergraduate group prior to the Kaggle challenge. As such, the choice of the undergraduate cohort could be inappropriate as a control group.
- To understand the assessment of student performance more thoroughly, it would be helpful if authors could provide the complete final exam at least for one of the two courses (perhaps in supplementary materials), clearly marking the questions pertaining to the Kaggle challenges.
- In section 3.1 authors introduce the metric that they use in the assessment of student performance. As authors discuss in the third paragraph of section 3.1, the constructed metric measures the consistency of students’ performance on a specific set of questions in relation to the overall exam performance. Subsequently authors use the defined metric for the assessment of performance, and attribute change in medians to improved performance (e.g. second paragraph of section 4.2). In reviewer’s view, change in medians shows change in consistency of performance, and that does not necessarily translate to improvement in learning. It is possible to construct a hypothetical example where the defined metric increases without any change in the performance on the regression- (or equivalently classification-) related part. It is not clear why authors used the “ratio” metric, as opposed to, for example, simply the percentage of the possible points received on a specific subset of questions (e.g.
regression- or classification-related). As such, the use of the proposed measure does not seem to be justified as being appropriate for the assessment of performance. If the currently used metric is retained, then the authors should explain the construction of it more clearly, exemplify it using numbers, and justify why an increase in the metric signals improvement in learning (rather than improvement in consistency).
- The entries in the “Median difference” column in Table 1 do not correspond to the differences in the median bars implied from Figure 1 (e.g. the difference between the median bars in the “regression” box from the left sub-figure and the “regression” box from the right sub-figure is not 0.1 as entered in Table 1).
Minor Comments
- It was not clear whether the training and testing sets noted in section 2.2 were the same across the two institutions.
- It is unclear if the survey mentioned in section 4.3 was anonymous. If it was, please stress that, and if it was not then address the bias that would be introduced in survey responses due to the lack of anonymity.
- It is mentioned in section 2.3 that there were 63 students randomized to regression and classification competitions, while in section 4.3 the total number goes down to 61. Please clarify.
- Further details behind Kaggle including but not limited to examples of popular datasets/competitions that it has hosted, data contributors, the motivation for contributing data, and the consequences of winning a competition or scoring high – would be helpful in the introduction.

Associate Editor
Comments to the Author:
I thought the article was well written and the topic definitely adds something new to the literature:

Notes:
1) There is no literature review. Even a short review of papers that have used projects or competitions would put this paper in context. Also some review (if it is out there) of your performance metric might help the readers to understand why you measured learning gains in this way.

2) Discussion of the logistics of setting up such a competition would be very helpful to the reader. The kaggle link should be more prominently placed in the article as well as your personal insight/experience to what kind of investment a professor needs to make in order to make this happen.

3) There needs to be a better explanation of why undergraduates were used as a control group at one of the universities. This jumped out to several reviewers.

4) I would argue for a data display/analysis that shows for each individual student the difference between their performance score on their "treatment" questions versus their "non-treatment" questions, i.e. the classification students that did “better than expected” on classification questions, how many of those also did “better than expected” on the regression questions and so on.

### 2018-12-20 submission1/response-to-review.txt

Reviewer(s)' Comments to Author:

Reviewer: 1

Comments to the Author
I enjoyed this paper and think it makes a contribution to the literature.  However,  I would like to see some improvements made.

Briefly discussing the experience from the professor side of things would be helpful.  So in addition to the student survey it would be helpful to read how hard it was to set up and work with the Kaggle competitions and any logistical challenges.

JULIA Setting up kaggle, and experience
DI Experience with kaggle in this class and others

Also more discussion of study limitations is needed.  For example,  in the ETC class the "treatment" is confounded with level of student ( graduate/ undergraduate).  Other limitations should be briefly discussed as well.  I also would like to see suggestions for future research.

DI on ETC2420

Sentence misworded in line 29 page 9

DI

Also in Figure 3, it would seem there is some non-linearity in a couple of the graphs.  I would like to see that mentioned and discussed.  It seems there is a point of diminishing return, in graphs E and F.

DI

@@

I also think it is a study weakness that all the survey questions are worded in a positive manner and this could be leading.  I realize given the study is done this can’t be changed but it would be worth noting in the discussion ways the survey might be improved.
It might be more productive in future work to directly ask for level of engagement first, then one could also ask students who did not do the competition as well the question.
“How would you rate your level of engagement in this course?”
Not at all engaged--- up to extremely engaged with choices in between.

JULIA

Reviewer: 2

Comments to the Author
Kaggle is an online platform that brings together two communities: one that is willing to contribute data and another that is interested in analyzing data. Kaggle has become one of the popular platforms used in quantitative statistical courses, where individual students or teams can obtain a data set to analyze, and often compete (for monetary prize or otherwise). The main aim of the article is to provide empirical evidence in support of the effectiveness of Kaggle competitions in improving student learning.

In terms of grammar, style, and narrative the article is written adequately. It gives a concise summary of the empirical results that it uses to support the claims of effectiveness. However, there are major issues related to the motivation and relatedly the (absence of) literature review; details and guidelines; and the quantitative assessment of effectiveness of the competition. The major issues are outlined below, followed by a few minor comments. In reviewer’s opinion, the adequate resolution of all of those observations is a necessary precondition for the publication of the work.

Major Comments
- The article does not provide a motivation for incorporating a predictive competition in a coursework. What is the educational aim of the competitive element as compared to more “standard” projects where students do not compete? Relatedly, what are the advantages and limitations of the proposed competitive format? A literature review, which currently is completely missing from the article, would be more than appropriate, summarizing thoroughly the existing work on project-based learning and putting the proposed format of Kaggle competitions vis-à-vis the project-based pedagogy.

JULIA lit review (same as reviewer 1)
DI motivation

@@

- In section 3.1 authors introduce the metric that they use in the assessment of student performance. As authors discuss in the third paragraph of section 3.1, the constructed metric measures the consistency of students’ performance on a specific set of questions in relation to the overall exam performance. Subsequently authors use the defined metric for the assessment of performance, and attribute change in medians to improved performance (e.g. second paragraph of section 4.2). In reviewer’s view, change in medians shows change in consistency of performance, and that does not necessarily translate to improvement in learning. It is possible to construct a hypothetical example where the defined metric increases without any change in the performance on the regression- (or equivalently classification-) related part. It is not clear why authors used the “ratio” metric, as opposed to, for example, simply the percentage of the possible points received on a specific subset of questions (e.g.
regression- or classification-related). As such, the use of the proposed measure does not seem to be justified as being appropriate for the assessment of performance. If the currently used metric is retained, then the authors should explain the construction of it more clearly, exemplify it using numbers, and justify why an increase in the metric signals improvement in learning (rather than improvement in consistency).

JULIA

- The entries in the “Median difference” column in Table 1 do not correspond to the differences in the median bars implied from Figure 1 (e.g. the difference between the median bars in the “regression” box from the left sub-figure and the “regression” box from the right sub-figure is not 0.1 as entered in Table 1).

DI

Minor Comments
- It was not clear whether the training and testing sets noted in section 2.2 were the same across the two institutions.

JULIA

### 2019-01-31 paper-kaggle.Rmd
reviewers comments addressing

@@

ETC2420/5242, titled Statistical Thinking, covers regression, and has a mix of undergraduate and postgraduate students. Only the 34 postgraduate (5242) students were required to participate in the kaggle competition, and competed in the regression (R) challenge. This was run independently from the University of Melbourne competition. The 145 undergraduate (2420) students were used as controls for examining performance of the postgraduate students. The competition ran for one month. Students formed their own teams of 2-4 members to compete. 
<!--
Several undergraduates also chose to compete individually. The material on regression methods, particularly the computational methods needed to successfully predict housing prices was new to both groups of students. 
-->
<!--
ETC3250, called Business Analytics, is an undergraduate course focusing on data mining. All students participated in a kaggle competition on the classification challenge. Because this group had no comparison group, it was difficult to assess performance. 
-->

@@

MAST90083 used \url{https://inclass.kaggle.com/c/XXX}. ETC2420/5242 used \url{https://inclass.kaggle.com/c/vitticeps}. 
-->

@@

@@

Figure \ref{fig:likert} shows the survey responses related to the kaggle competition, for MAST90083 and ETC5242. The response rate for MAST90083 was 55%, with 34 of 61 students completed the survey. The response rate for ETC5242 was 50%, 17 students out of 34 completed the survey. Overwhelmingly, students reported that they found the competition interesting and helpful for their learning in the course.

>>>>

The training and the testing data sets of the Melbourne auction price data were similar but not identical across the two institutions. Some of the variables in the data set were simulated, e.g. property land size and house size. The simulated data was generated slightly differently for different institutions. This was done deliberately to prevent students passing answers from one institution to another.

Students received clean data ready for modelling as we only aimed to examine students ability to apply covered material. If the instructor's aim is rather to mimic real-life situation when prior to modelling/forecasting data needs to be processed and cleaned, then this task should be left to students. In our opinion, this approach is more suitable for advanced classes.

@@

The reason for this strategy was first to motivate each of the students to think about modelling and be actively engaged in the competitions through individual submission. The lecturer allowed to create groups towards the end of the competition to illustrate the advantages of group work and mixing models. Another motivation for this stratagem was the university policy, requiring a strategy to assign students individually in group assignments.

0ne of the 63 students didn't took part in the competition. Another student didn't sit the exam. Therefore, the analysed sample size reduced to 61.

ETC2420/5242, titled Statistical Thinking, covers regression, and has a mix of undergraduate and postgraduate students. Only the 34 postgraduate (5242) students were required to participate in the kaggle competition, and competed in the regression (R) challenge. This was run independently from the University of AB competition. The 145 undergraduate (2420) students were used as controls for examining performance of the postgraduate students. The competition ran for one month. Students formed their own teams of 2-4 members to compete. 
<!--
Several undergraduates also chose to compete individually. The material on regression methods, particularly the computational methods needed to successfully predict housing prices was new to both groups of students. 
-->
<!--
ETC3250, called Business Analytics, is an undergraduate course focusing on data mining. All students participated in a kaggle competition on the classification challenge. Because this group had no comparison group, it was difficult to assess performance. 
-->

@@

MAST90083 used \url{https://www.kaggle.com/c/amiaspam} and \url{https://www.kaggle.com/c/melbprice}. ETC2420/5242 used \url{https://inclass.kaggle.com/c/vitticeps}. 
-->

## Working with the Kaggle InClass platform

Creating a new competition is surprisingly easy. 
Information on setting up a Kaggle InClass challenge can be found at https://www.kaggle.com/about/inclass/overview. 
To ease the setting process, we created a short video illustrating the steps to establish a competition....

Once you open the competition, it runs without any intervention from the instructor. The instructor can monitor students progress: the number of submissions, student scores and even the uploaded data at any time. When the competition ends the "Leaderboard" page provides a list of students ordered by the final score. It also provides all the scores from all past submissions (under "Raw Data" on "Public Leaderboard").

From our experience, we recommend limiting the number of daily submissions to one. It encourages students to think about more efficient improvement of their model before the next submissions.  We also believe that it is better to start the competition in the second half of the semester and finish it two to one week before the end of the semester. It allows to cover many different statistical methods that students may apply during the competition. Finishing some time before the end of the semester reduces students anxiety at the end of the semester and the approaching exams.

@@

```{r crossperformances, echo=FALSE, fig.cap="\\label{fig:crossperformances} Students performance in classification and regression questions by competition type. Each quarter of the plot also contains  the counted number of students in the brackets. The number of students that participated in the classification competition is shown first, followed by the number of students that took part in the  regression competition. ", fig.width=6,fig.height=5  }

t <- UniMelbData %>%
 count(Proj_alloc, perf_class >=1, perf_reg >=1, sort = TRUE) %>%
  setNames(., c("Proj_alloc", "outperf_class", "outperf_reg" , "n"))

a <- paste( "(", as.character( filter(t, Proj_alloc == "classification" , outperf_class == FALSE,  outperf_reg == FALSE)["n"]), ",", 
                 as.character( filter(t, Proj_alloc == "regression" ,     outperf_class == FALSE,  outperf_reg == FALSE)["n"]) ,")")

b <- paste( "(", as.character( filter(t, Proj_alloc == "classification" , outperf_class == FALSE,  outperf_reg == TRUE)["n"]), ",", 
                 as.character( filter(t, Proj_alloc == "regression" ,     outperf_class == FALSE,  outperf_reg == TRUE)["n"]) ,")")

c <- paste( "(", as.character( filter(t, Proj_alloc == "classification" , outperf_class == TRUE,  outperf_reg == TRUE)["n"]), ",", 
                 as.character( filter(t, Proj_alloc == "regression" ,     outperf_class == TRUE,  outperf_reg == TRUE)["n"]) ,")")

d <- paste( "(", as.character( filter(t, Proj_alloc == "classification" , outperf_class == TRUE,  outperf_reg == FALSE)["n"]), ",", 
                 as.character( filter(t, Proj_alloc == "regression" ,     outperf_class == TRUE,  outperf_reg == FALSE)["n"]) ,")")

# filter(UniMelbData, Proj_alloc == "regression" , perf_class < 1, perf_reg >=1) %>% select(., perf_class, perf_reg)
# filter(UniMelbData, Proj_alloc == "regression" , perf_class < 1, perf_reg <1) %>% select(., perf_class, perf_reg)

ggplot(UniMelbData, aes(x=perf_class, y=perf_reg, color = Proj_alloc)) +
      geom_point() +
      ylab("Regression questions performance") + 
      xlab("Classification questions performance") +
      theme(legend.position="bottom") +
      xlim(0,2) + ylim(0, 2) +
      scale_color_discrete(name="Competiotion", 
                          breaks=c("regression", "classification"),   
                          labels=c("Regression", "Classification")) +
      geom_hline(yintercept = 1, color = "gray", size=0.7) +  #linetype="dotted",
      geom_vline(xintercept = 1, color = "gray", size=0.7) +
      annotate("text", x = c(0.1,0.1,1.9,1.9), y=c(0.1,1.9,1.9,0.1), 
               label = c(  a,  #(1,1)
                           b,  # "(2,2)"
                           c,  #"(3,3)"
                           d )) #"(4,4)"
      # annotate("text", x=0.1, y=1.8, label= "(,)")

Figure \ref{fig: crossperformances} presents students' scores for classification and regression questions. A score over 1 is considered outperformance (relative to the expectation). Quarters one and three include students that underperform or outperform on both types of questions, respectively. In both cases the number of students that participated in the classification competition is very close to the number of students that participated in the regression competition (excluding a few regression students on the border of score 1). Students in quarters two and four outperform on one type of questions but not on the other type. We can see that more regression students outperform on regression questions than classification students (12 vs. 7). Similarly, classification students do better on classification questions (11 vs. 3). This is another evidence towards positive influence of the data competition on student's performances.

@@

Figure \ref{fig:likert} shows the survey responses related to the kaggle competition, for MAST90083 and ETC5242. The response rate for MAST90083 was 55%, with 34 of 61 students completing the survey. 
The response rate for ETC5242 was 50%, 17 students out of 34 completed the survey. Overwhelmingly, students reported that they found the competition interesting and helpful for their learning in the course.

After collecting the survey from the students we were told that the questions about students engagement were positively worded. This may have potentially led to some bias. An improved way could be to ask directly about student's engagement, e.g. "How would you rate your level of engagement in this course?" with set answer options of "not at all engaged" -- up to "extremely engaged" with several choices in between. Another improvement could be asking ETC2420 students that didn't take part in the competition about their level of engagement and compare the answers with other students of ETC5242. We acknowledge that the differences in the engagement levels may not necessarily be a result of participation in the competition but it is still an interesting aspect.

The survey was not anonymous. However, the results became available to the lectures only after all the grades were realised to the students. This point was emphasized in the instructions to the students at the beginning of the survey.
